Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
fixed emptyset

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model Algorithm}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} $r_1, r_2$: parameters controlling the learning rate 
    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign \{\}$
    \For{$i \assign 0$ \textbf{to} $\minit$}
    \State $\sigma \assign \text{generate uniformly at random}$
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor
    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S,\rho)$\label{line:uborda}\Comment{Learning Step}
    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
   % \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     \State $\sigma \assign$ sample a permutation from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    % \Comment{Generate $\beta$ permutations}
    %\State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
    %\Statex \Comment{Select the farthest from $S$}
   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor 
    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

@@

\myparagraph{UMM}
%
UMM uses a budget of evaluations ($\FEmax$) of  $\minit=10$ solutions sampled uniformly.
The other two parameters of UMM are $r_1$ and $r_2$ that determine the learning rate $\rho$ in uBorda, and consequently, determine the
balance between exploration and exploitation. In the experimental section we
will study their impact.

We consider 3 different values for the maximum budget of $\FEmax=100, 200,400$ evaluations of the
actual objective function. In a white-box context, state-of-the-art algorithms
for the LOP and PFSP typically evaluate thousands of solutions,
thus, the budget considered here for the black-box context is extremely
limited.

>>>>

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model Algorithm}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} $r_1, r_2$: parameters controlling the learning rate 
    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign \emptyset$
    \For{$i \assign 0$ \textbf{to} $\minit$}
    \State $\sigma \assign \text{generate uniformly at random}$
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor
    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S,\rho)$\label{line:uborda}\Comment{Learning Step}
    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
   % \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     \State $\sigma \assign$ sample a permutation from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    % \Comment{Generate $\beta$ permutations}
    %\State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
    %\Statex \Comment{Select the farthest from $S$}
   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor 
    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

@@

\myparagraph{UMM}
%
UMM initially samples $\minit=10$ solutions uniformly.
The other two parameters of UMM are $r_1$ and $r_2$ that determine the learning rate $\rho$ in uBorda, and consequently, determine the
balance between exploration and exploitation. In the experimental section we
will study their impact.

We consider three different values $\FEmax \in\{100, 200,400\}$ for the maximum budget of  evaluations of the
actual objective function. In a white-box context, state-of-the-art algorithms
for the LOP and PFSP typically evaluate thousands of solutions,
thus, the budget considered here for the black-box context is extremely
limited.
